---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# scHiCSRS

<!-- badges: start -->
<!-- badges: end -->

The goal of scHiCSRS is to impute single cell HiC data using a self representation smoothing approach.

## Installation

The scHiCSRS package has the following R-package dependencies: SAVER, keras, tensorflow, Rtsne, ggplot2, ggpubr, and mclust. The dependent packages will be automatically installed along with scHiCSRS. You can use the following commands to install scHiCSRS from GitHub.

``` r
# Install and load "devtools" package. 
install.packages("devtools")
library("devtools")

# Install "scHiCSRS" package from github.
install_github("Queen0044/scHiCSRS")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(scHiCSRS)
## data("simudat")
#simudat_res=SRS(simudat, windowsize=2, nbins=61, epochs = 100, estimates.only = FALSE)
```

For more information of functions, please read the vignettes.
